---
title: "Statistics to support Sander *et al.*"
author: "Liz Sander"
date: "17 May 2016"
output: html_document
---

```{r session_setup, echo=FALSE, warning = FALSE, message = FALSE}
source('ImbalanceTableSampling.R')
source('CalculateEvenness.R')
source('MutualInformation.R')
library(dplyr)
library(agricolae) ## for HSD test
```

```{r DC_Imbalance_ttest, echo=FALSE, warning = FALSE, message = FALSE}
## paired test for the difference between degree-corrected
## and non-degree-corrected imbalance values.
## Separate tests performed with and without concomitant
## predation.
## Function taken from ImbalanceTableSampling.R
out <- ImbalanceTableSampling('../../Results/ImbalanceSampling')
## format output for paired t-test, pairing degree corrected/uncorrected
imbTable <- out %>% unite(tmp, Imbalance, probimbal) %>%
    tidyr::spread(key = DC, value = tmp) %>%
    tidyr::separate(Corrected,
                    c('Corrected_Imbalance_c', 'Corrected_probimbal_c'),
                    sep = '_') %>%
    tidyr::separate(Uncorrected,
                    c('Uncorrected_Imbalance_c', 'Uncorrected_probimbal_c'),
                    sep = '_') %>%
    dplyr::mutate(Corrected_Imbalance = as.numeric(Corrected_Imbalance_c)) %>%
    dplyr::mutate(Corrected_probimbal = as.numeric(Corrected_probimbal_c)) %>%
    dplyr::mutate(Uncorrected_Imbalance = as.numeric(Uncorrected_Imbalance_c)) %>%
    dplyr::mutate(Uncorrected_probimbal = as.numeric(Uncorrected_probimbal_c)) %>%
    dplyr::select(-Corrected_Imbalance_c, -Corrected_probimbal_c,
                  -Uncorrected_Imbalance_c, -Uncorrected_probimbal_c)
## split into concomitant/no concomitant and run separate t-tests
imbTablePar <- imbTable %>% dplyr::filter(parcon == 'No_Concomitant')
imbTableParCon <- imbTable %>% dplyr::filter(parcon == 'Concomitant')
imbDiffPar <- t.test(imbTablePar$Uncorrected_Imbalance,
                     imbTablePar$Corrected_Imbalance,
                     alternative = "two.sided", paired = TRUE)
imbDiffParCon <- t.test(imbTableParCon$Uncorrected_Imbalance,
                        imbTableParCon$Corrected_Imbalance,
                        alternative = "two.sided", paired = TRUE)
```

The estimated difference between uncorrected and corrected imbalance
values across all webs is `r as.vector(imbDiffPar$estimate)`
(p = `r imbDiffPar$p.value`) when concomitant predation is excluded,
and
`r as.vector(imbDiffParCon$estimate)` (p = `r imbDiffParCon$p.value`)
when it is included.

```{r concom_difference_ttest, echo = FALSE, warning = FALSE, message = FALSE}
## format output for paired t-test, pairing concomitant/no concomitant
imbTableConcom <- out %>% unite(tmp, Imbalance, probimbal) %>%
    tidyr::spread(key = parcon, value = tmp) %>%
    tidyr::separate(No_Concomitant,
                    c('Par_Imbalance_c', 'Par_probimbal_c'), sep = '_') %>%
    tidyr::separate(Concomitant,
                    c('ParCon_Imbalance_c', 'ParCon_probimbal_c'), sep = '_') %>%
    dplyr::mutate(Par_Imbalance = as.numeric(Par_Imbalance_c)) %>%
    dplyr::mutate(Par_probimbal = as.numeric(Par_probimbal_c)) %>%
    dplyr::mutate(ParCon_Imbalance = as.numeric(ParCon_Imbalance_c)) %>%
    dplyr::mutate(ParCon_probimbal = as.numeric(ParCon_probimbal_c)) %>%
    dplyr::select(-Par_Imbalance_c, -Par_probimbal_c,
                  -ParCon_Imbalance_c, -ParCon_probimbal_c)
## split into degree corrected/uncorrected and run separate t-tests
imbTableDC0 <- imbTableConcom %>% dplyr::filter(DC == 'Uncorrected')
imbTableDC1 <- imbTableConcom %>% dplyr::filter(DC == 'Corrected')
imbDiffDC0 <- t.test(imbTableDC0$ParCon_Imbalance,
                     imbTableDC0$Par_Imbalance,
                     alternative = "two.sided", paired = TRUE)
imbDiffDC1 <- t.test(imbTableDC1$ParCon_Imbalance,
                     imbTableDC1$Par_Imbalance,
                     alternative = "two.sided", paired = TRUE)
```

The estimated difference in imbalance between a web without
concomitant predation and a web with it is
`r as.vector(imbDiffDC0$estimate)` (p = `r imbDiffDC0$p.value`) when
not correcting for degree, and `r as.vector(imbDiffDC1$estimate)` (p =
`r imbDiffDC1$p.value`) using the degree-corrected model.

```{r degree_difference_ANOVA, echo = FALSE, warning = FALSE, message = FALSE}
## ANOVAs to test for differences in indegree and outdegree across
## ecological categorizations. Separate ANOVAs for in and out degree,
## and for with/without concomitant predppation.
## Code snippets taken from DegreeHist.R
## Note that this is a one-way ANOVA, so the unbalanced design
## is not a major issue.
webs <- c('BahiaFalsa', 'Carpinteria', 'Flensburg', 'Otago',
          'PuntaBanda', 'Sylt', 'Ythan')
pc <- c('Par', 'ParCon')
degreedf <- data.frame(Web = character(0),
                       parcon = character(0),
                       Degree = numeric(0),
                       InOut = character(0),
                       Ecological_Group = character(0))
for(web in webs){
    for(pcname in pc){
        if(pcname == 'Par') dfpc <- 'No Concomitant' else dfpc <- 'Concomitant'
        mat <- as.matrix(read.table(
            paste0('../../Data/Dunne2013-', web, '-',
                   pcname, '-mat.txt')))
        classvec <- as.vector(as.matrix(read.table(
            paste0('../../Data/Dunne2013-', web, '-Category.txt'))))
        tmp <- data.frame(Web = rep(web, nrow(mat)*2),
                          parcon = rep(dfpc, nrow(mat)*2),
                          Degree = c(rowSums(mat), colSums(mat)),
                          InOut = c(rep('Indegree', nrow(mat)),
                              rep('Outdegree', nrow(mat))),
                          Ecological_Group = rep(classvec, 2))
        degreedf <- degreedf %>% dplyr::bind_rows(tmp)
    }
}
## ANOVA tests
aovInPar <- aov(Degree ~ Ecological_Group,
                data = filter(degreedf, InOut == 'Indegree', parcon == 'No Concomitant'))
HSDaovInPar <- agricolae::HSD.test(aovInPar, "Ecological_Group", group = TRUE)
aovOutPar <- aov(Degree ~ Ecological_Group,
                 data = filter(degreedf, InOut == 'Outdegree',
                     parcon == 'No Concomitant'))
HSDaovOutPar <- agricolae::HSD.test(aovOutPar, "Ecological_Group", group = TRUE)
aovInParCon <- aov(Degree ~ Ecological_Group,
                   data = filter(degreedf, InOut == 'Indegree', parcon == 'Concomitant'))
HSDaovInParCon <- agricolae::HSD.test(aovInParCon, "Ecological_Group", group = TRUE)
aovOutParCon <- aov(Degree ~ Ecological_Group,
                data = filter(degreedf, InOut == 'Outdegree', parcon == 'Concomitant'))
HSDaovOutParCon <- agricolae::HSD.test(aovOutParCon, "Ecological_Group", group = TRUE)
```

Degree differs significantly across ecological categorizations both
with (indegree: p = `r summary(aovInPar)[[1]][[1,"Pr(>F)"]]`,
outdegree: `r summary(aovOutPar)[[1]][[1,"Pr(>F)"]]`) and without
(indegree: `r summary(aovInParCon)[[1]][[1,"Pr(>F)"]]`, outdegree:
`r summary(aovOutParCon)[[1]][[1,"Pr(>F)"]]`) concomitant
predation. Differences in means were summarized by calculating the
Tukey's HSD:

Indegree, no concomitant:
```{r echo=FALSE}
HSDaovInPar$groups
```

Outdegree, no concomitant:
```{r echo=FALSE}
HSDaovOutPar$groups
```

Indegree, with concomitant:
```{r echo=FALSE}
HSDaovInParCon$groups
```

Outdegree, with concomitant:
```{r echo=FALSE}
HSDaovOutParCon$groups
```

```{r calculate_evenness, echo=FALSE, warning = FALSE, message = FALSE}
## from CalculateEvenness.R
evennessDF <- BuildEvennessDF('../../Results/GroupModel/')
## split by concomitant predation and run t-tests
evenPar <- dplyr::filter(evennessDF, parcon == 'Par')
evenParCon <- dplyr::filter(evennessDF, parcon == 'ParCon')
ttestPar <- t.test(evenPar$Uncorrected, evenPar$Corrected,
                   alternative = 'two.sided', paired = TRUE)
ttestParCon <- t.test(evenParCon$Uncorrected, evenParCon$Corrected,
                      alternative = 'two.sided', paired = TRUE)
```

Uncorrected partitions are significantly less even across webs both
for webs with (estimate: `r ttestPar$estimate`, p =
`r ttestPar$p.value`) and without (estimate: `r ttestParCon$estimate`, p
= `r ttestParCon$p.value`) concomitant predation.
